---
title: "Morning vs Night"
author: "Richard Sprague"
date: "8/25/2022"
output:
  github_document: default
  html_document: default
---

What happens to my mouth microbiome throughout the day?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require("treemap")) install.packages("treemap")
library(tidyverse)
source("bristle_read.R")

```

## Baseline

This sample was taken in June on a random, typical morning first thing after waking. 

```{r}
baseline_sample <- read_bristle_table(filepath=file.path("data", "BristleHealthRaw.xlsx"))

baseline_sample %>% treemap::treemap(dtf=., index = c("genus","species"),
                 vSize="abundance",
                 type = "categorical",
                 vColor = "genus",
                 position.legend = "none",
                 title = "My Mouth Microbes (Baseline)")
```


## Morning

This sample was collected a few minutes after waking up. I usually wear an orthodontic retainer, so this is my oral microbiome *with* the retainer.

```{r}

am_sample <- read_bristle_table(filepath=file.path("data", "Bristle-2022-08-02-AM.xlsx"))


am_sample %>% 
treemap::treemap(dtf=., index = c("genus","species"),
                 vSize="abundance",
                 type = "categorical",
                 vColor = "genus",
                 position.legend = "none",
                 title = "My Mouth Microbes (Morning)")

```



## Evening


I went the rest of the day as normal. This sample was collected at 10pm, about three hours after my last meal, and before inserting the retainer or brushing my teeth.

```{r}

pm_sample <- read_bristle_table(filepath=file.path("data", "Bristle-2022-08-02-PM.xlsx"))

pm_sample %>% 
treemap::treemap(dtf=., index = c("genus","species"),
                 vSize="abundance",
                 type = "categorical",
                 vColor = "genus",
                 position.legend = "none",
                 title = "My Mouth Microbes (Evening)")
```

