---
title: "Bristle and Phyloseq"
author: "Richard Sprague"
date: "`r Sys.Date()`"
output:
  github_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(phyloseq)
library(psmr)  # use library function instead.
if(!require("bristler")) devtools::install_github("personalscience/bristler")


```


## Read Bristle sample data

```{r}

bristle_data <-bristler::read_bristle_table(filepath=file.path("data/BristleHealthRaw.xlsx"))

```

## Create a mapfile


Experiment dataframe with the following exact column names (additional columns are okay but will be ignored):
  "ssr": sequencing revision (e.g. 42578).  Your column probably repeats many SSRs.
  "tax_name": (e.g. "Bifidobacterium").
  "count" : reads for that taxa at that SSR (e.g. 2137)

Mapping file: a dataframe that contains:
  "ssr": exact same SSRs as experiment dataframe above
  attributes columns: as many as you like.  (e.g. "yogurteater", "geo", "kefireater", "gender")

```{r}

bristle_data$ssr <- 1234
mapfile <- tibble(ssr=1234,label="baseline")

```

## Create an experiment file

```{r}
bristle_experiment <- bristle_data %>% transmute(ssr,
                                                 tax_name = genus,
                                                 count = abundance * 10000,
                                                 tax_rank = "genus")
```

## Make the Phyloseq object

```{r}

e.ps <- bristler::phyloseqize(bristle_data, mapfile = mapfile)

```


```{r}
plot_bar(e.ps, fill = "Genus")
plot_heatmap(e.ps, method = "CCA", distance = "none",fill = "Genus")

```

Now attempt other Phyloseq operations.
```{r}

estimate_richness(e.ps, measures = c("InvSimpson","Shannon"))

plot_richness(e.ps, measures = c("InvSimpson","Shannon"))


```

